---
title: "Assignment2_q3"
author: "Soham Sengupta"
date: "15/11/2022"
Roll: MDS202241
---

## Problem 3: Analysis of `faithful` datasets.

Consider the `faithful` datasets:
```{r}
attach(faithful)
hist(faithful$waiting,xlab = 'waiting',probability = T,col='pink',main='')
```

Fit following three models using MLE method and calculate **Akaike information criterion** (aka., AIC) for each fitted model. Based on AIC decides which model is the best model? Based on the best model calculate the following probability
$$
\mathbb{P}(60<\texttt{waiting}<70)
$$

(i) **Model 1**:
$$
f(x)=p*Gamma(x|\alpha,\sigma_1)+(1-p)N(x|\mu,\sigma_2^2),~~0<p<1
$$
(ii) **Model 2**:
$$
f(x)=p*Gamma(x|\alpha_1,\sigma_1)+(1-p)Gamma(x|\alpha_2,\sigma_2),~~0<p<1
$$

(iii) **Model 3**:
$$
f(x)=p*logNormal(x|\mu_1,\sigma_1^2)+(1-p)logNormal(x|\mu_1,\sigma_1^2),~~0<p<1
$$
```{r}
library(tidyverse)
library(scales)
x <- sort(waiting)
```
```{r}
##model1
p = length(x[x<65])/length(x)
m1 = mean(x[x<65])
var1 = var(x[x<65])
m2 = mean(x[x>=65])
sd2 = sd(x[x>=65])
theta_initial = c(p, m1^2/var1, var1/m1, m2, sd2)
negativeLogLikelihood <- function(theta, x){
  p = theta[1]
  alpha1 = theta[2]
  sigma1 = theta[3]
  meu = theta[4]
  sigma2 = theta[5]
  l = sum(log(p*dgamma(x, shape = alpha1, scale = sigma1) + (1-p)*dnorm(x, mean = meu, sd = sigma2)))
  return(-l)
}
fit = optim(theta_initial, negativeLogLikelihood, x=x, lower = c(0,0,0,-Inf, 0), upper = c(1, Inf, Inf, Inf, Inf), method = "L-BFGS-B")
theta_hat1 = fit$par
theta_hat1

p = theta_hat1[1]
alpha1 = theta_hat1[2]
sigma1 = theta_hat1[3]
meu = theta_hat1[4]
sigma2 = theta_hat1[5]

model1 = p*dgamma(x, shape = alpha1, scale = sigma1) + (1-p)*dnorm(x, mean = meu, sd = sigma2)
aic1 = 2*5 + negativeLogLikelihood(theta_hat1, x)
sprintf("AIC for model 1: %f",aic1)
```
```{r}
##model 2
p = length(x[x<65])/length(x)
m1 = mean(x[x<65])
var1 = var(x[x<65])
m2 = mean(x[x>=65])
var2 = var(x[x>=65])
theta_initial = c(p, m1^2/var1, var1/m1, m2^2/var2, var2/m2)
negativeLogLikelihood2 <- function(theta, x){
  p = theta[1]
  alpha1 = theta[2]
  sigma1 = theta[3]
  alpha2 = theta[4]
  sigma2 = theta[5]
  l = sum(log(p*dgamma(x, shape = alpha1, scale = sigma1) + (1-p)*dgamma(x, shape = alpha2, scale = sigma2)))
  return(-l)
}
fit = optim(theta_initial, negativeLogLikelihood2, x=x, lower = c(0,0,0,0,0), upper = c(1, Inf, Inf, Inf, Inf), method = "L-BFGS-B")
theta_hat2 = fit$par
theta_hat2

p = theta_hat2[1]
alpha1 = theta_hat2[2]
sigma1 = theta_hat2[3]
alpha2 = theta_hat2[4]
sigma2 = theta_hat2[5]

model2 = p*dgamma(x, shape = alpha1, scale = sigma1) + (1-p)*dgamma(x, shape = alpha2, scale = sigma2)
aic2 = 2*5 + negativeLogLikelihood2(theta_hat2, x)
sprintf("AIC for model 2: %f",aic2)
```
```{r}
p = length(x[x<65])/length(x)
m1 <- mean(x[x<65])
v1 <- var(x[x<65])
sigma1 <- log((v1/m1^2) + 1)
meu1 <- log(m1) - sigma1/2
m2 <- mean(x[x>=65])
v2 <- var(x[x>=65])
sigma2 <- log((v2/m2^2) + 1)
meu2 <- log(m2) - sigma2/2
theta_initial = c(p, meu1, sqrt(sigma1), meu2, sqrt(sigma2))
negativeLogLikelihood3 <- function(theta, x){
  p = theta[1]
  meu1 = theta[2]
  sigma1 = theta[3]
  meu2 = theta[4]
  sigma2 = theta[5]
  l = sum(log(p*dlnorm(x, meanlog = meu1, sdlog = sigma1) + (1-p)*dlnorm(x, meanlog = meu2, sdlog = sigma2)))
  return(-l)
}
fit = optim(theta_initial, negativeLogLikelihood3, x=x, lower = c(0,-Inf,0,-Inf,0), upper = c(1, Inf, Inf, Inf, Inf), method = "L-BFGS-B")
theta_hat3 = fit$par
theta_hat3

p = theta_hat3[1]
meu1 = theta_hat3[2]
sigma1 = theta_hat3[3]
meu2 = theta_hat3[4]
sigma2 = theta_hat3[5]

model3 = p*dlnorm(x, meanlog = meu1, sdlog = sigma1) + (1-p)*dlnorm(x, meanlog = meu2, sdlog = sigma2)
aic3 = 2*5 + negativeLogLikelihood3(theta_hat3, x)
sprintf("AIC for model 3: %f",aic3)
```
```{r}
hist(faithful$waiting,xlab = 'waiting',probability = T,col='pink',main='')
lines(x, model1, col = 'red')
lines(x, model2, col = 'green')
lines(x, model3, col = 'blue')
print("The AIC of all the three models")
result <- data.frame(models = c("Gamma + Normal", "Gamma + Gamma", "Lognormal + Lognormal"), AIC = c(aic1, aic2, aic3))
result
```
Therefore, from the above output we see that the AIC of Model 3 is lowest. Hence it has the best fit among the three model. Hence we choose the lognormal-lognormal model for getting the probability.

```{r}
prob = p*plnorm(70, meanlog = meu1, sdlog = sigma1) + (1-p)*plnorm(70, meanlog = meu2, sdlog = sigma2)-p*plnorm(60, meanlog = meu1, sdlog = sigma1) - (1-p)*plnorm(60, meanlog = meu2, sdlog = sigma2)
sprintf("Hence based on the best fit model P(60 < waiting < 70) = %f",prob)
```


