---
title: "Assignment2_q5"
author: "Soham Sengupta"
date: "15/11/2022"
output:
  html_document:
    df_print: paged
Roll: MDS202241
---

## Problem 5: Computational Finance - Modelling Stock prices

Following piece of code download the prices of TCS since 2007

```{r}
library(quantmod)
getSymbols('TCS.NS')
tail(TCS.NS)
```
Plot the adjusted close prices of TCS
```{r}
plot(TCS.NS$TCS.NS.Adjusted)
```
**Download the data of market index Nifty50**. The Nifty 50 index indicates how the over all market has done over the similar period.
```{r}
getSymbols('^NSEI')
tail(NSEI)
```
Plot the adjusted close value of Nifty50
```{r}
plot(NSEI$NSEI.Adjusted)
```
### Log-Return 
We calculate the daily log-return, where log-return is defined as
$$
r_t=\log(P_t)-\log(P_{t-1})=\Delta \log(P_t),
$$
where $P_t$ is the closing price of the stock on $t^{th}$ day.

```{r}
TCS_rt = diff(log(TCS.NS$TCS.NS.Adjusted))
Nifty_rt = diff(log(NSEI$NSEI.Adjusted))
retrn = cbind.xts(TCS_rt,Nifty_rt) 
retrn = na.omit(data.frame(retrn))

plot(retrn$NSEI.Adjusted,retrn$TCS.NS.Adjusted
     ,pch=20
     ,xlab='Market Return'
     ,ylab='TCS Return'
     ,xlim=c(-0.18,0.18)
     ,ylim=c(-0.18,0.18))
grid(col='grey',lty=1)
```
+ Consider the following model:

$$
r_{t}^{TCS}=\alpha + \beta r_{t}^{Nifty} + \varepsilon,
$$
where $\mathbb{E}(\varepsilon)=0$ and $\mathbb{V}ar(\varepsilon)=\sigma^2$.

1. Estimate the parameters of the models $\theta=(\alpha,\beta,\sigma)$ using the method of moments type plug-in estimator discussed in the class.

2. Estimate the parameters using the `lm` built-in function of `R`. Note that `lm` using the OLS method.

3. Fill-up the following table

Parameters | Method of Moments | OLS
-----------|-------------------|-----
$\alpha$   |     0.000463      |0.0004628
$\beta$    |     0.743684      |0.7436838
$\sigma$   |     0.016185      |0.01619

4. If the current value of Nifty is 18000 and it goes up to 18200. The current value of TCS is Rs. 3200/-. How much you can expect TCS price to go up?
```{r}
##Part 1
##Using method of moments in simple linear regression model we have the following eqn
##1->E[Y-alpha-beta*X]=0
##2->E[(Y-alpha-beta*X)*X]=0
##3->E[e^2]=sigma^2
##Solving we get
##beta=cov(x,y)/var(x)
## alpha= E(Y) - beta*E(x)
## sigma=

beta= cov(retrn$NSEI.Adjusted,retrn$TCS.NS.Adjusted)/var(retrn$NSEI.Adjusted)
alpha= mean(retrn$TCS.NS.Adjusted) - beta* mean(retrn$NSEI.Adjusted)
yhat= alpha+beta*retrn$NSEI.Adjusted
sigma=(var(retrn$TCS.NS.Adjusted-yhat))^0.5
sprintf("Estimated value of Alpha: %f",alpha)
sprintf("Estimated value of Beta: %f",beta)
sprintf("Estimated value of sigma: %f",sigma)

```
```{r}
## Part 2
simple_fit = lm(retrn$TCS.NS.Adjusted~retrn$NSEI.Adjusted, data=retrn)
summary(simple_fit)
```
```{r}
##Part 4
r4n=log(18200)-log(18000)
r4t=0.0004628+0.7436838*r4n
x=exp(r4t+log(3200))
sprintf("Expected Price of TCS: %f",x)

```