---
title: "Problem 1"
author: "Ujan Dasgupta"
date: "15/11/2022"
output: html_document
---
1. 

Given,  
\begin{equation}
\begin{alignedat}{3}
\mathbb{P}[X = 0] = a\\
\mathbb{P}[X = x] = ar^x
\end{alignedat}
\end{equation}
where $x\geq 0$ and $x\in \mathbb{N}$. Now we know that
\begin{equation}
\begin{alignedat}{3}
& \sum_{x = 0}^\infty \mathbb{P}[X=x] = 1\\
\Rightarrow\; & \sum_{x = 0}^\infty ar^x = 1\\
\Rightarrow\; & a\left[1 + r + r^2 + \cdots\right] = 1\\
\Rightarrow\; & \frac{a}{1-r} = 1\\
\Rightarrow\; & a = 1-r
\end{alignedat}
\end{equation}

Since we must have $0<a<1$ we should have $0<r<1$. This is our necessary condition. The probability distribution model is as follows:
\begin{equation}
\begin{alignedat}{3}
p(x) &= ar^x = (1-r)r^x, \text{ when } x>0, x\in \mathbb{N}\\
&= 0, \text{ otherwise}
\end{alignedat}
\end{equation}

2. 

For the expectation and variance to exist, $\mathbb{E}[\mid X\mid]$ and $\mathbb{E}\left[\left|X^2\right|\right]$ must exist. Here
\begin{equation}
\begin{alignedat}{3}
\mathbb{E}[\mid X\mid] &= \sum_{x = 0}^\infty \mid x\mid (1-r)r^x\\
                       &= \sum_{x = 0}^\infty x(1-r)r^x\\
                       &< \infty
\end{alignedat}
\end{equation}
$\therefore\;\; \mathbb{E}\left[\left|X\right|\right]$ exists.
\begin{equation}
\begin{alignedat}{3}
\mathbb{E}[\mid X\mid] &= \sum_{x = 0}^\infty \mid x\mid (1-r)r^x\\
                       &= \sum_{x = 0}^\infty x(1-r)r^x\\
                       &< \infty
\end{alignedat}
\end{equation}

3.
We now have: \begin{align*} \mathbb{E}(X) &= \sum_{i} i\mathbb{P}(X=i)
\\ &= \sum_{i} i(1-r)r^i \\ \implies \frac{\mathbb{E}(X)}{1-r} &=
\sum_{i} ir^i \\ \implies \frac{\mathbb{E}(X)}{1-r} &= \frac{r}{(1-r)^2}
\\ \implies \mathbb{E}(X) &= \frac{r}{1-r} \end{align*}

Observe that \begin{align*} \mathbb{E}(X^2) &= \sum_{i} i^2
\mathbb{P}(X=i) \\ &= \sum_{i} i^2(1-r)r^i \end{align*}

Calculating the sum, we obtain: $$\mathbb{E}(X^2) =
\frac{r(r+1)}{(1-r)^2}$$

To calculate the variance, now: \begin{align*} \text{Var}(X) &=
\mathbb{E}(X^2) - [\mathbb{E}(X)]^2 \\ &= \frac{r(r+1)}{(1-r)^2} -
\frac{r^2}{(1-r)^2} \\ &= \frac{r}{(1-r)^2} \end{align*}

4.
From historical data we have Mean = 1.5,
Hence, using the above model we have E(x)=1.5
or, r/(1-r) = 1.5
or, r= 1.5 - 1.5r
```{r}
r=1.5/2.5
sprintf("Estimated value of r = %f",r)
 ##Part a
##P(Home team will have atleast 1 goal) = 1- P(X=0)
p_4a=1-(1-r)
sprintf("P(Home team will have atleast 1 goal) = %f",p_4a)

##Part b
#P(Home team will have atleast 1 goal but less than 4 goals)=P(X=1)+P(X=2)+P(X=3)
p_4b= (1-r)*(r+r*r+r*r*r)
sprintf("P(Home team will have atleast 1 goalbut less than 4 goals) = %f",p_4b)

```

5.
We now have to use Poisson Distribution, from the historical data we have mean=1.5
```{r}
##Part A
##P(Home team will have atleast 1 goal) = 1- P(X=0)
p_5a=1-dpois(0,1.5)
sprintf("P(Home team will have atleast 1 goal) = %f",p_5a)

##Part b
#P(Home team will have atleast 1 goal but less than 4 goals)=P(X=1)+P(X=2)+P(X=3)
p_5b= dpois(1,1.5) + dpois(2,1.5) + dpois(3,1.5)
sprintf("P(Home team will have atleast 1 goal but less than 4 goals) = %f",p_5b)

```

6.
I will prefer model 1 (Geometric model) over the poisson probability model.
As historical data suggests Var(x)>E(x)
which is a case in geometric distribution as seen in part 3,
whereas, in case of Poisson Distribution E(X) = Var(X), which doesn't hold for
this data.


7.   
Write down the likelihood functions of your newly defined probability models and Poisson models. Clearly mention all the assumptions that you are making.     

For the Poisson distribution: $$L(\lambda|\mathbf{x}) = \prod_{i=1}^{n}
\frac{e^{-\lambda}\lambda^{x_i}}{x_i!} = e^{-n\lambda}
\prod_{i=1}^{n}\frac{\lambda^{x_i}}{x_i!}$$

For the geometric distribution: $$L(r|\mathbf{x}) = \prod_{i=1}^{n}
(1-r)r^{x_i} = (1-r)^n\,r^{\sum_{i=1}^{n} x_i}$$